Model

### 2019-09-16 stats/kids_combination.Rmd
fix dob on comb data file

## importing data
# files <- dir("~/Work/Spin/raw_data/kids/combination/")
# 
# raw_data <- data_frame()
# for (f in files) {
#   jf <- paste("~/Work/Spin/raw_data/kids/combination/",f,sep="")
#   jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
#   date <- str_sub(jf,56,str_length(jf)-11)
#   id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
#   raw_data <- bind_rows(raw_data, id)
# }
# 
# d_raw <- raw_data %>%
#   mutate(discourseNovel = str_sub(discourseNovel,11),
#          mutExNovel = str_sub(mutExNovel, 10),
#          subid = ifelse(subid == "190726_6_come", "190726_6_com", subid))
# 
# d_raw %>%
#   filter(subid == "190727_1_com")
# 
# 
# log <- read_excel("../../SPIN-subject_log.xlsx", 1)%>%
#   filter(Condition == "com")%>%
#   mutate(dob = as.character(dob),
# #         dob = ifelse(subid == "190313_7_me", "2015-11-15",dob),
#          dob = as.Date(dob))%>%
#   select(subid,experimenter,keep_drop,sex,dob)
# 
#  
# d_raw%>%
#   group_by(subid)%>%
#   summarise(n())%>%
#   filter(!(subid %in% unique(log$subid)))
# 
#  
# d <- left_join(d_raw,log, by = "subid")%>%
#   filter(keep_drop == "keep")%>%
#   mutate(age_num = lubridate::time_length(difftime(test_date,dob), "years"),
#          check_age = ifelse(substr(age_num,1,1) == subage, T, F))
# 
# 
# write_csv(d, "../data/combination.csv")

>>>>

# importing data
files <- dir("~/Work/Spin/raw_data/kids/combination/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/Spin/raw_data/kids/combination/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,56,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

d_raw <- raw_data %>%
  mutate(discourseNovel = str_sub(discourseNovel,11),
         mutExNovel = str_sub(mutExNovel, 10),
         subid = ifelse(subid == "190726_6_come", "190726_6_com", subid))

d_raw %>%
  filter(subid == "190727_1_com")

log <- read_excel("../../SPIN-subject_log.xlsx", 1)%>%
  filter(Condition == "com")%>%
  mutate(dob = as.character(dob),
         dob = ifelse(subid == "190620_7_com", "2016-06-20",dob),
         dob = ifelse(subid == "190620_8_com", "2014-10-04",dob),
         dob = ifelse(subid == "190620_14_com", "2016-12-16",dob),
         dob = ifelse(subid == "190711_4_com", "2017-12-11",dob),
         dob = ifelse(subid == "190516_03_com", "2016-05-16",dob),
         dob = ifelse(subid == "190711_4_com", "2016-12-11",dob),

         dob = as.Date(dob))%>%
  select(subid,experimenter,keep_drop,sex,dob)

d_raw%>%
  group_by(subid)%>%
  summarise(n())%>%
  filter(!(subid %in% unique(log$subid)))

d <- left_join(d_raw,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))

d%>%
  filter(check_age == F)%>%
  select(subid, subage,test_date,dob,age_num,check_age)

write_csv(d, "../data/combination.csv")

### 2019-09-16 stats/model.Rmd
fix dob on comb data file

```{r}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)

@@

#### WORK IN PROGRESS: nesting within participant

@@

  foreach(function(row){
      var prior = [row.prior, 1-row.prior]
      foreach(function(item){
        var itemLineParameters = all_item_parameters[item]
        var sem_knowledge = logistic(itemLineParameters.intercept + 
            itemLineParameters.slope * row.age)
        var speakerOptimality = speakerOptimalityParameters.intercept  + speakerOptimalityParameters.slope * row.age

@@

    //  query.add(["prediction", item, row.age, row.alignment], ["NA", Math.exp(modelPredictions.score(1))])

      }, familiars)
   }, dataFromR.comb_data)

@@

me_model_data <- read_csv("../data/me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(age = age_num - min(age_num),
         nov_int = novel_params$Intercept,
         nov_slope = novel_params$age_num,
         prior = plogis(nov_int + age * nov_slope))

comb_model_data <- bind_rows(
  me_model_data %>%
    distinct(age, prior) %>%
    mutate(prior = 1 - prior, 
         alignment = "incongruent"),

   me_model_data %>%
    distinct(age, prior) %>%
    mutate(alignment = "congruent")
)

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = comb_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH",
                        samples = 100000,
                        burn = 50000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 5000)
)

# run variants
# start with variant 3
# 3 cores, 3 chains 
# see how long it takes and then run over night

#saveRDS(sem_know_model, "../saves/sem_know_model_variant2_100k.rds")

@@

@@

ggplot(comb_model_pred_agg, aes(x=age,y = mean, col = item, fill = item))+
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = .5, colour = NA)+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Combination Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  facet_grid(Chain~alignment)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()+
  scale_fill_viridis_d()

@@

>>>>

```{r}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)
library(matrixStats)

@@

# Nesting within participant

```{r}
novel_bm <- readRDS("../saves/novelty_model.rds")

data_comb_log <- read_csv("../data/combination.csv") %>%
  filter(condition != "train")%>%
  mutate(alignment = condition,
         item = familiarObject, 
         age = age_num - (min(age_num))) %>%
  mutate(nov_int = novel_params$Intercept,
         nov_slope = novel_params$age_num,
         prior = ifelse(alignment == "congruent", plogis(nov_int + nov_slope* age),1-plogis(nov_int + age * nov_slope)))

data_comb_log

## Pragmatics model

@@

var all_log_likes = map(function(row) { 
      var prior = [row.prior, 1-row.prior]
        var itemLineParameters = all_item_parameters[row.item]
        var sem_knowledge = logistic(itemLineParameters.intercept + 
            itemLineParameters.slope * row.age)
        var speakerOptimality = speakerOptimalityParameters.intercept  + speakerOptimalityParameters.slope * row.age

@@

        return modelPredictions.score(row.correct)

    //  query.add(["prediction", item, row.age, row.alignment], ["NA", Math.exp(modelPredictions.score(1))])
  // return loglike
   }, dataFromR.comb_data)

  var log_likelihood = sum(all_log_likes)
  query.add(["log_like", "NA", "NA"], ["NA", log_likelihood])

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH",
                        samples = 100000,
                        burn = 50000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 1000)
)

#saveRDS(sem_know_model, "../saves/sem_know_model_variant3_100k_log_like.rds")

@@

log_like<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "item","age","alignment"), sep = ",")%>%
  filter(parameter == "log_like")%>%
  separate(value, into = c("na", "log_like"),sep = ",")%>%
  select(-na, -item, -age, -alignment)%>%
  mutate(log_like =as.numeric(str_remove(log_like,"[)]")))%>%
  group_by(Chain)%>%
  summarize(logP = logSumExp(log_like))

## Flat prior model

```{r work in progress}
flat_prior <- '
var data = dataFromR.data

var all_log_likes = map(function(row) { 
      var prior = [row.prior, 1-row.prior]
        var itemLineParameters = all_item_parameters[row.item]
        var sem_knowledge = logistic(itemLineParameters.intercept + 
            itemLineParameters.slope * row.age)
        var speakerOptimality = speakerOptimalityParameters.intercept  + speakerOptimalityParameters.slope * row.age

        var modelPredictions = pragmaticListener({label: "novel_word"}, [.5,.5], speakerOptimality, sem_knowledge)

    //    display(modelPredictions.score(row.correct))

        return modelPredictions.score(row.correct)

    //  query.add(["prediction", item, row.age, row.alignment], ["NA", Math.exp(modelPredictions.score(1))])
  // return loglike
   }, dataFromR.comb_data)

  var log_likelihood = sum(all_log_likes)
  query.add(["log_like", "NA", "NA"], ["NA", log_likelihood])

sem_know_model_flat_prior<- webppl(
  program_code = paste(rsaUtils, rsaModel, flat_prior , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH",
                        samples = 100000,
                        burn = 50000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 1000)
)

saveRDS(sem_know_model_flat_prior, "../saves/sem_know_model_variant3_100k_flat_prior_log_like.rds")

model_pred<- sem_know_model_flat_prior %>%
  separate(Parameter, into = c("parameter", "item","age","alignment"), sep = ",")%>%
  filter(parameter == "prediction")%>%
  separate(value, into = c("na", "me_effect"),sep = ",")%>%
  select(-na)%>%
  mutate(age = as.numeric(age),
         me_effect =as.numeric(str_remove(me_effect,"[)]")))

log_like_flat_prior<- sem_know_model_flat_prior %>%
  separate(Parameter, into = c("parameter", "item","age","alignment"), sep = ",")%>%
  filter(parameter == "log_like")%>%
  separate(value, into = c("na", "log_like"),sep = ",")%>%
  select(-na, -item, -age, -alignment)%>%
  mutate(log_like =as.numeric(str_remove(log_like,"[)]")))%>%
  group_by(Chain)%>%
  summarize(logP = logSumExp(log_like))

@@

ggplot(comb_model_pred_agg, aes(x=age,y = mean, col = item, fill = item))+
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = .5, colour = NA)+
  geom_smooth(data = data_comb, aes(x = age_num, y = correct, col), col = "black", fill = "grey",  method = "loess", se = F, alpha = .2, span = 2)+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Combination Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  facet_grid(Chain ~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()+
  scale_fill_viridis_d()

@@

```{r}
ggsave("../graphs/model_pred_data.pdf", width = 20, height = 8, scale = 1.2)
```
